---
title: "Informedb Software Quality Metrics"
author: "Liam Whitenack"
date: "`r Sys.Date()`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    dev: pdf
    highlight: tango
  html_document:
    theme: default
    self_contained: true
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    smart: true
    dev: svg
---
# Setup
```{r setup, include = FALSE}
# Set knitr options
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	fig.align = "center",
	fig.asp = 0.618,
	fig.width = 5,
	cache = FALSE,
	dpi = 120,
	out.width = "70%"
)
# Load required packages
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(infer))
suppressPackageStartupMessages(library(modelr))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(rlang))
suppressPackageStartupMessages(library(plotly))
Jira <- read_csv("Informedb Jira Data.csv")
```
# Tidy Data
## Create the resolved column
```{r resolved_date}
# This code block updates the "Resolved" column.
# The resolved column only contains a value if the row's value in the Status
# column is "Resolved". 
# If the logical statement is TRUE then the Resolved column's value is the same
# as the Updated column.
for(i in 1:length(Jira$Status)){
  Jira$Resolved[i]<-NA
}
for(i in 1:length(Jira$Status)){
        if(Jira$Status[i] == "Resolved"){
          Jira$Resolved[i]<-Jira$Updated[i]
        }
}
```
## Rename columns for computation
```{r mutate_Jira}
Jira <- Jira %>%
    mutate(
      Priority = recode(
       Priority,
#          `1 - Trivial` = "R4",
          `Trivial` = "1 - Trivial",
#          `2 - Minor` = "R4",
          `Minor` = "2 - Minor",
#          `3 - Major` = "R3",
          `Major` = "3 - Major",
#          `4 - Critical` = "R2",
          `Critical` = "4 - Crtitical",
#          `5 - Blocker` = "R1",
          `Blocker` = "5 - Blocker",
      ), 
      # Rename each value to R1, R2, R3, and R4, 
      # the military's preferred terms.
      Assignee_dev = recode(
       Assignee,
          `DanielBiester` = TRUE,
          `DavidTuell` = TRUE,
          `DrewPolhamus` = TRUE,
          `JosephWallace` = TRUE,
          `MasonVines` = TRUE,
          `MattChild` = TRUE,
          `NancyKraus` = TRUE,
          `PeterMorrissey` = TRUE,
          `PeteSchatz` = TRUE,
          `ScottSummerford` = TRUE,
          .default = FALSE
      ), 
      Reporter_dev = recode(
       Reporter,
          `DanielBiester` = TRUE,
          `DavidTuell` = TRUE,
          `DrewPolhamus` = TRUE,
          `JosephWallace` = TRUE,
          `MasonVines` = TRUE,
          `MattChild` = TRUE,
          `NancyKraus` = TRUE,
          `PeterMorrissey` = TRUE,
          `PeteSchatz` = TRUE,
          `ScottSummerford` = TRUE,
          .default = FALSE
      ),
      completed = recode(
        Status,
        Resolved = TRUE,
        Done = FALSE,
        Closed = FALSE,
        `In Progress` = FALSE,
        `Ready For Review` = FALSE,
        `Ready For Test` = FALSE,
        Reopened = FALSE,
        .default = FALSE
      ), # set the value of each resolved issue to 1 for computational purposes
      open = recode(
        Status,
        Resolved = FALSE,
        Done = TRUE,
        Closed = FALSE,
        `In Progress` = TRUE,
        `Ready For Review` = TRUE,
        `Ready For Test` = TRUE,
        Reopened = TRUE,
        Todo = TRUE,
        .default = TRUE
      ), # set the value of each resolved issue to 1 for computational purposes
      failed_fix = recode(
        Status,
        Resolved = FALSE,
        Done = FALSE,
        Closed = TRUE,
        `In Progress` = FALSE,
        `Ready For Review` = FALSE,
        `Ready For Test` = FALSE,
        Reopened = FALSE,
        .default = FALSE
      ), # set the value of each resolved issue to 1 for computational purposes
      Parent = `Parent id`,
      Issue = `Issue id`,
      Type = `Issue Type`,
)
```
## Create new useful columns from old ones
```{r calculations}
Jira <- Jira %>%
    mutate(
      date_created = as.Date(strptime(Created, format = "%m/%d/%Y %H:%M")),
      date_updated = as.Date(strptime(Updated, format = "%m/%d/%Y %H:%M")),
      date_resolved = as.Date(strptime(Resolved, format = "%m/%d/%Y %H:%M")), 
      # read columns as dates instead of characters
      time_spent = date_updated - date_created, 
      # create the time_spent parameter, which is computed by finding 
      # the difference between the date of creation and last day updated
      time_spent_no_outliers = time_spent,
      # create a row with a new name to be written later
      repair_time = date_resolved - date_created,
      # create the repair time parameter, which is computed by finding 
      # the difference between the date of creation and day resolved
      repair_time_no_outliers = repair_time,
      # create a row with a new name to be written later
    )
```
## Make a two time columns without outliers
### Create repair time and time spent IQRs
```{r IQR, echo=FALSE}
Q1ts <- quantile(Jira$time_spent, .25)
Q3ts <- quantile(Jira$time_spent, .75)
Q1rt <- quantile(Jira$repair_time, na.rm = TRUE, .25)
Q3rt <- quantile(Jira$repair_time, na.rm = TRUE, .75)
```
### Get rid of the cells that are outliers
```{r no_outliers, echo=FALSE}
for(i in 1:length(Jira$Priority)){
  if((Jira$time_spent[i] < Q1ts) || (Jira$time_spent[i] > Q3ts)){
    Jira$time_spent_no_outliers[i] <- NA
  }
  else{
    Jira$time_spent_no_outliers[i] <- Jira$time_spent[i]
  }
  if(is.na(Jira$repair_time[i]) == FALSE){
    if(((Jira$repair_time[i] < Q1rt) || (Jira$repair_time[i] > Q3rt))){
      Jira$repair_time_no_outliers[i] <- NA
    }
    else{
      Jira$repair_time_no_outliers[i] <- Jira$repair_time[i]
    }
  }
}
```
## Select only the columns being used
```{r Jira_reduced}
# Select only the rows I need
Jira_reduced <- Jira %>%
  select(
    Priority, # The importance of fixing each bug R1 is the highest importance,
    # R4 is the lowest
    `Issue Type`,
    Status, # the status of each progress i.e ""In progress",
    # "todo", or "resolved"
    Assignee, # The person assigned to the issue.
    # Most issues do not typically have an assignee
    Reporter, #Typically the same as the creator  
    date_created, # The day each issue was first created
    date_updated, # The last time the issue was edited
    date_resolved, # The last time the issue was edited
    # IF the status is "resolved"
    time_spent, # date updated - date created
    time_spent_no_outliers, # time_spent without outliers
    repair_time, # date updated - date created IF the status is "resolved"
    repair_time_no_outliers, # repair_time without outliers
    completed, # TRUE if completed, FALSE if not
    open, # TRUE if open, FALSE if not
    failed_fix, # TRUE if "closed", FALSE if not
    Assignee_dev, # TRUE if the project was assigned to a dev
    Reporter_dev, # TRUE if reported by a dev
  )
```
## make a column that indicates whether the bug was created in the last 28 days
```{r index_28}
# This code block will do the same thing as before,
# but uses the last 28 days instead of 365.
# Because this is being used for summary statistics and not a graph,
# this block will record a "TRUE" or "FALSE" for each row
# to be later added onto the tibble
n <- 0
counter <- 0
while(n < 1){
  counter <- counter + 1
  if(Jira_reduced$date_created[counter] > (max(Jira_reduced$date_created)-28)){
    index_28 <- TRUE
    n = 1
    counter = counter + 1
  }
  else{
    index_28 <- FALSE
    n = 1
    counter = counter + 1
  }
}
for(i in counter:length(Jira_reduced$date_created)){
  if(Jira_reduced$date_created[i] > (max(Jira_reduced$date_created)-28)){
    index_28 <- c(index_28, TRUE)
  }
  else{
    index_28 <- c(index_28, FALSE)
  }
}
```
```{r index_28}
# This code block will do the same thing as before,
# but uses the last 28 days instead of 365.
# Because this is being used for summary statistics and not a graph,
# this block will record a "TRUE" or "FALSE" for each row
# to be later added onto the tibble
n <- 0
while(n < 1){
  counter <- counter + 1
  if(is.na(Jira_reduced$Assignee[i])){
    index_a <- FALSE
    n = 1
  }
  else{
    index_a <- TRUE
    n = 1
  }
}
for(i in 2:length(Jira_reduced$date_created)){
  if(is.na(Jira_reduced$Assignee[i])){
    index_a <- c(index_a, FALSE)
  }
  else{
    index_a <- c(index_a, TRUE)
  }
}
```
```{r last28}
# add a new row to Jira_reduced that has a value of TRUE
# if the issue was created in the last 28 days,
# FALSE if the issue was created before
Jira_reduced <- Jira_reduced %>%
  mutate(
    last28 = index_28,
    assigned = index_a
  )
```
## Rename the ugly columns
```{r muatate_reduced}
Jira_reduced <- Jira_reduced %>%
  mutate(
    Issue_Type = `Issue Type`,
    `Date Created` = date_created,
    `Date Updated` = date_updated,
    `Date Resolved` = date_resolved,
    `Time Spent` = time_spent,
    `Time Spent (Without Outliers)` = time_spent_no_outliers,
    `Repair Time` = repair_time,
    `Repair Time (Without Outliers)` = repair_time_no_outliers,
    `Assigned to a Developer` = Assignee_dev,
    `Reported by a Developer` = Reporter_dev
  )
```
## Make the completed neat dataset
```{r Jira_tidy_select}
Jira_tidy <- Jira_reduced %>%
  select(
    Priority,
    `Issue_Type`,
    Status,
    Assignee,
    Reporter,
    `Date Created`,
    `Date Updated`,
    `Date Resolved`,
    `Time Spent`,
    `Time Spent (Without Outliers)`,
    `Repair Time`,
    `Repair Time (Without Outliers)`,
    completed,
    open,
    failed_fix,
    assigned,
    last28,
    `Assigned to a Developer`,
    `Reported by a Developer`
  )
```
# Create a Dataset for the Military Presentation
## List every column marked with "bug"
```{r bug_index}
# record the row number of each row

#this while loop initializes the array and sets its first value
n <- 0
counter <- 0
while(n < 1){
  counter <- counter + 1
  if(Jira$`Issue Type`[counter] == "Bug" || Jira$`Issue Type`[counter]
     == "Pre-Release Bug"){
    bug_index <- counter
    n = 1
    counter = counter + 1
  }
}
# this for loop checks to see if the issues have the Issue Type "Bug".
# If they do, mark their row number for later
for(i in counter:length(Jira$`Issue Type`)){
  if(Jira$`Issue Type`[i] == "Bug" || Jira$`Issue Type`[i] == "Pre-Release Bug"){
    bug_index <- c(bug_index, i)
  }
}
```
## Rename Priorities
```{r Priority_rename}
# This code block's function is to rename all of the Priorities to the military's preferred terms.

Jira_tidy <- Jira_tidy %>%
  mutate(
      Priority = recode(
       Priority,
          `1 - Trivial` = "R4",
          `Trivial` = "R4'",
          `2 - Minor` = "R4",
          `Minor` = "R4",
          `3 - Major` = "R3",
          `Major` = "R3",
          `4 - Critical` = "R2",
          `Critical` = "R2",
          `5 - Blocker` = "R1",
          `Blocker` = "R1",
      )
)
```
## Create new dataset with only bugs
```{r Jira_bugs}
# make a new tibble with only the rows recorded in bug_index
Jira_bugs <- Jira_tidy[c(bug_index),]
```
## Make Reliability Metrics
```{r Summary_Statistics}
# Calculate the summary statistics by their Priority
summary_stats <- Jira_bugs %>%
  group_by(Priority) %>%
  summarise( 
    `Defects in the last four weeks` = sum(last28), 
    # count the number of rows that were created in the last 28 days
    MTTR = (mean((`Repair Time`), na.rm = TRUE)), 
    # Repair time average
    DRE = sum(completed) / n(), 
    # count the number of rows marked as completed
    Open = sum(open) 
    # count the number of rows that aren't closed or resolved
  ) 
```
# Output
```{r write_csv}
write_csv(Jira_tidy, "Informedb_tidy.csv")
write_csv(Jira_bugs, "Informedb_NSWCCD_data.csv")
write_csv(summary_stats, "Informedb_Reliability_Metrics.csv")
```

